---
title: "Accessing UFO sightings data"
author: "Colleen Nell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Accessing UFO sightings data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package currently provides 2 functions to gather data on UFO sightings from the National UFO Report Center (www.nuforc.org). These functions take dates and return sightings data. 

To install this package:  
```{r, eval=F}
library(devtools)
library(ggplot2)

install_github("collnell/Xfiles")
library(Xfiles)

```   

```{r, echo=F}
library(Xfiles)
library(ggplot2)

```  

```{r, include=F}
theme_UFO<-function(base_size = 18, base_family = "Helvetica", legend.location = "right"){
  theme(
    line =               element_line(colour = "green", size = 1, linetype = 1, 
                                      lineend = "round"),
    rect =               element_rect(fill = "white", colour = "white", size = 0.5, linetype = 1),
    text =               element_text(family = base_family, face = "plain", colour = "white",
                                      size = base_size, hjust = 0.5, vjust = 0.5, angle = 0,
                                      lineheight = 0.9, margin=margin(), debug=FALSE),
    axis.text =          element_text(size = rel(0.8), colour = "white"),
    strip.text =         element_text(size = rel(0.8)),
    
    axis.line.x =        element_line(color = "white", size=1),
    axis.line.y =        element_line(color = "white", size=1),
    axis.text.x =        element_text(face = "plain", margin=margin(2,0,0,0)),
    axis.text.y =        element_text(face = "plain", margin=margin(0,2,0,0)),
    axis.ticks =         element_line(colour = "white", size= .6),
    axis.title =         element_text(colour = "white", face = "bold"),
    axis.title.x =       element_text(margin=margin(10,0,0,0)),
    axis.title.y =       element_text(angle = 90, margin=margin(0,10,0,0)),
    axis.ticks.length =  unit(0.15, "cm"),
    
    legend.background =  element_rect(colour = NA),
    legend.key =         element_blank(),
    legend.key.size =    unit(1.2, "lines"),
    legend.key.height =  NULL,
    legend.key.width =   NULL,
    legend.text =        element_text(size = rel(0.8)),
    legend.text.align =  NULL,
    legend.title =       element_text(size = rel(0.8), face = "bold", hjust = 0),
    legend.title.align = NULL,
    legend.position =    legend.location,
    legend.direction =   NULL,
    legend.justification = "center",
    legend.box =         NULL,
    
    panel.background =   element_rect(fill = "black", colour = NA),
    panel.border =       element_blank(),
    panel.grid.major =   element_blank(),
    panel.grid.minor =   element_blank(),
    
    strip.background =   element_rect(fill = "black", size = 0.5),
    strip.text.x =       element_text(),
    strip.text.y =       element_text(angle = -90),
    
    plot.background =    element_rect(colour = "black", fill="black"),
    plot.title =         element_text(size = rel(1.2)),
    plot.margin =        unit(c(1, 1, 1, 1), "lines"),
    
    complete = TRUE
  )
    
}

UFOcount <- function(date=NA){
  url_month<-"http://www.nuforc.org/webreports/ndxevent.html"##all the data if there, want to pull only for the month give
  lines<-readLines(url_month)
  date_pattern<-'<FONT style=FONT-SIZE:11pt FACE=\"Calibri\" COLOR=#000000>([^<]*)</TD>'
  datalines<-grep(date_pattern,lines[26:length(lines)], value=TRUE)
  
  getexpr = function(s,g)substring(s,g,g+attr(g,'match.length')-1)
  gg = gregexpr(date_pattern,datalines)
  matches = mapply(getexpr,datalines,gg)
  result = gsub(date_pattern,'\\1',matches)
  names(result) = NULL##UFO counts by month
  
  datpat<-'<FONT style=FONT-SIZE:11pt FACE=\"Calibri\" COLOR=#000000><A HREF= ndxe([^<]*)</A></TD>'
  datadate<-grep(datpat,lines[26:length(lines)], value=TRUE)
  ggd = gregexpr(datpat,datadate)
  matchesd = mapply(getexpr,datadate,ggd)
  resultdate = gsub(datpat,'\\1',matchesd)
  names(resultdate) = NULL
  ##remove every 201611.html>
  ufos_month<-data.frame(Date = sapply(strsplit(resultdate, ">", fixed=TRUE),"[[",2),
                         UFOs = result)##make new data frame with UFO sightings by month
  ufos_month$UFOs<-as.integer(as.character(ufos_month$UFOs))
  ufos_month$Date<-as.Date(as.yearmon(as.character(ufos_month$Date), "%m/%Y"), format="%m/%d/%Y")
  if (!is.na(date)) {
    ufos_month<-ufos_month%>%
      filter(ufos_month$Date == as.Date(paste0(date,"-01")))
    return(ufos_month)
  } else{
  return(ufos_month)
  }
  
}

UFOobs <- function(year=NA, month=NA, year_end= NA, month_end = NA) {
  if (is.na(year_end) | is.na(month_end)){
    if (nchar(month)<2){
      month<-paste0(0,as.integer(month))
    }else{
      month<-month
    }
    url <- paste0("http://www.nuforc.org/webreports/ndxe", year, month, ".html", sep="")
    resp<-htmltab(url, which=1, encoding = "UTF-8")
    resp$Date<-as.Date(resp$`Date / Time`, format="%m/%d/%y %H:%M")
    resp$Month<-month(resp$Date)
    resp$Year<-year(resp$Date)
    resp<-resp%>%
      select(-Posted, -`Date / Time`)
    return(resp)
    
  } else {
    date_start <- as.Date(paste0(year,"-",month,"-01"))
    date_end <- as.Date(paste0(year_end,"-",month_end,"-01"))
    
    mos<-seq(ymd(date_start),ymd(date_end), by="month")
    months<-data.frame(dates=mos)
    months$y=as.integer(year(months$dates))
    months$m<-ifelse(nchar(as.integer(month(months$dates)))<2, 
                     paste0(0,as.integer(month(months$dates))),as.integer(month(months$dates)))
    mot<-paste0(months$y,months$m)
    
    functest<-function(mot){
      url <- paste0("http://www.nuforc.org/webreports/ndxe", mot, ".html", sep="")
      resp<-htmltab(url, which=1, encoding = "UTF-8")
      return(resp)
    }
    
    output<-do.call(rbind, lapply(mot, functest))
    output$Date<-as.Date(output$`Date / Time`, format="%m/%d/%y %H:%M")
    output$Month<-month(output$Date)
    output$Year<-year(output$Date)
    output<-output%>%
      select(-Posted, -`Date / Time`)
    return(output)
  }
}

```
## UFOcount()  
This function returns a dataframe containing the total count of UFO observations for a given month & year. Dates need to be entered in a 'YYYY-mm' format

```{r}
UFOcount(date='2016-11')

```  
If no date is provided, by default all monthly observations on record are returned by date.  
```{r}
df<-UFOcount()
str(df)

```  
The UFO records contain 891 month records. This includes data from WAY back, which might not be very reliable. Consistent monthly records on UFO sightings begin around 1950.  
```{r}
df<-filter(df, df$Date>=as.Date('1950-01-01'))
str(df)

```  
The theme_UFO() theme for ggplot2 may come in handy in plotting sightings data:  

UFO monthly sightings since 1950  
```{r,fig.width=7, fig.height=5} 
ggplot(df, aes(x=Date, y=UFOs))+geom_point(size=1, shape=8, color="green")+theme_UFO()
```
```{r, fig.width=7, fig.height=5}
df$year<-as.integer(year(df$Date))
df$month<-as.integer(month(df$Date))
df<-top_n(df,25)

ggplot(df, aes(x=month, y=UFOs, group=year, color=year))+geom_line()+theme_UFO()+scale_color_continuous(low="greenyellow", high="green4")

```


##UFOobs()  
This function gathers higher resolution data on each UFO sighting for a given month and year. Data returned include: Date, City, State, Shape, and a Summary. 

```{r}
df<-UFOobs(year=1980, month=01)
str(df)

```  
Observation data can also be drawn for a date range. Warning: this is a larger dataset and may take a while if requesting a large date range. It is adviseable to use to UFOcount() function to consider the number of observations in the data first.  

```{r, eval=F}
df<-UFOobs(year=2014, month=01, year_end= 2016, month_end = 12)

se <- function(x) sd(x)/sqrt(length(x))

df<-df%>%
  select(Date, State, year)%>%
  group_by(State, year)%>%
  summarize(UFOs = length(Date))%>%
  group_by(State)%>%
  summarize(UFO_mean = mean(UFOs), UFO_se = se(UFOs))

```  

```{r, include=F}
df<-read.csv("/Users/colleennell/Dropbox/Projects/Xfiles/UFOs_2000s.csv")

```

```{r, fig.width=7, fig.height=5}

df<-top_n(df, 20)
##average number of UFO sightings per year by state
ggplot(df, aes(x=reorder(State,UFO_mean), y=UFO_mean))+
  geom_bar(stat="identity", color="green", fill="black")+
  geom_errorbar(aes(ymin=UFO_mean - UFO_se, ymax=UFO_mean + UFO_se), width=.1, color="white")+
  labs(x="State", y="Average Annual UFO sightings")+
  theme_UFO()


```


  
  
  
  