---
title: "Accessing UFO sightings data"
author: "Colleen Nell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Accessing UFO sightings data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package currently provides 2 functions to gather data on UFO sightings from the National UFO Report Center (www.nuforc.org). These functions take dates and return sightings data. 

To install this package:  
```{r, eval=F}
install_github("collnell/Xfiles")
library(Xfiles)

```  
```{r, echo=F}
library(Xfiles)

```  
## UFOcount()  
This function returns a dataframe containing the total count of UFO observations for a given month & year. Dates need to be entered in a 'YYYY-mm' format

```{r}
UFOcount(date='2016-11')

```  
If no date is provided, by default all monthly observations on record are returned by date.  
```{r}
df<-UFOcount()
str(df)

```  
The UFO records contain 891 month records. This includes data from WAY back, which might not be very reliable. Consistent monthly records on UFO sightings begin around 1950.  
```{r}
df<-filter(df, df$Date>=as.Date('1950-01-01'))
str(df)

```  
The theme_UFO() theme for ggplot2 may come in handy in plotting sightings data:  

UFO monthly sightings since 1950  
```{r} 
ggplot(df, aes(x=Date, y=UFOs))+geom_point(size=1, shape=8, color="green")+theme_UFO()
```

  
  
  
  